---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# CBC bindings for R

This package provides bindings to the [COIN-CBC solver](https://projects.coin-or.org/Cbc).

It is currently work in progress. I have only tested it on a Mac so far.

## Quickstart

* `-lCbc` needs to be in your `PKG_LIBS` flags.
* The header files of the coin libraries (`cbc`, `clp`, `osi`, `coinutils`) should to be in the include path of your cpp compiler.

```{r, eval=FALSE}
devtools::install_github("dirkschumacher/rcbc")
```

```{r}
library(rcbc)
# max 1 * x + 2 * y
# s.t.
#   x + y <= 1
#   x, y integer
A <- matrix(c(1, 1), ncol = 2, nrow = 1)
result <- CBC_solve(
 obj = c(1, 2),
 mat = A, # <- can also be a sparse matrix
 is_integer = c(TRUE, TRUE),
 row_lb = -Inf, row_ub = 1, max = TRUE)
result
```


## TODO

* Easy installation on all platforms
* Parallel solving
* Add configuration for cut generators
* Add callback support
* Write a ROI plugin.

## Contribution

Feel free to open issues and send PRs.
