---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# CBC bindings for R
<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/Lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html)
[![check-Ubuntu](https://img.shields.io/github/actions/workflow/status/dirkschumacher/rcbc/check-ubuntu.yaml?branch=master&label=Ubuntu)](https://github.com/dirkschumacher/rcbc/actions)
[![check-Windows](https://img.shields.io/github/actions/workflow/status/dirkschumacher/rcbc/check-windows.yaml?branch=master&label=Windows)](https://github.com/dirkschumacher/rcbc/actions)
[![check-macOS](https://img.shields.io/github/actions/workflow/status/dirkschumacher/rcbc/check-macos.yaml?branch=master&label=macOS)](https://github.com/dirkschumacher/rcbc/actions)
[![Documentation](https://img.shields.io/github/actions/workflow/status/dirkschumacher/rcbc/documentation.yaml?branch=master&label=Documentation)](https://github.com/dirkschumacher/rcbc/actions)
[![Coverage Status](https://img.shields.io/codecov/c/github/dirkschumacher/rcbc?label=Coverage)](https://codecov.io/github/dirkschumacher/rcbc?branch=master)
[![CRAN status](https://www.r-pkg.org/badges/version/rcbc)](https://CRAN.R-project.org/package=rcbc)
<!-- badges: end -->

The _rcbc_ package provides an interface to the [_CBC_ (COIN-OR branch and cut)](https://github.com/coin-or/Cbc) solver. Specifically, _CBC_ is an open-source mixed integer programming solver that is developed as part of the [Computational Infrastructure for Operations Research (COIN-OR) project](https://www.coin-or.org/). By interfacing with the _CBC_ solver, the _rcbc_ package can be used to generate optimal solutions to optimization problems. Please note that this package is under active development and is still a work in progress.

## Installation

The package is not yet available on [The Comprehensive R Archive Network](https://cran.r-project.org/). To install this package, please use the following _R_ code to install it from the [source code repository on GitHub](https://github.com/dirkschumacher/rcbc). Please note that [CBC solver](https://github.com/coin-or/Cbc) header and library files also need be installed prior to installing this _R_ package (see below for details).

```{r, eval = FALSE}
if (!require(remotes))
  install.packages("remotes")
remotes::install_github("dirkschumacher/rcbc")
```

### Windows

The package can be installed from source when the [Rtools](https://cran.r-project.org/bin/windows/Rtools/) software is installed. Specifically, the [CBC solver](https://github.com/coin-or/Cbc) header and library files are automatically downloaded from [RWinLib](https://github.com/rwinlib/cbc).

### Linux

#### Debian/Ubuntu

The following system command can be used to install dependences.

```
sudo apt-get install coinor-libcbc-dev coinor-libclp-dev
```

#### Fedora

The following system command can be used to install dependencies.

```
sudo yum install coin-or-Cbc-devel coin-or-Clp-devel
```

### macOS

The following system command can be used to install dependencies using [Homebrew package manager](https://brew.sh/). After installing CBC and its dependencies, they need to linked in order to install the _rcbc_ package. **Please note that if you have previously installed these software, then they will be overwritten with the newer versions.**

```
brew tap coin-or-tools/coinor
brew install coin-or-tools/coinor/cbc
brew link cbc --force
brew link coinutils --force
brew link osi --force
brew link clp --force
brew link cgl --force
```

## Usage

Here we will provide a brief example showing how the package can be used to solve an optimization problem (see [package vignette](https://dirkschumacher.github.io/rcbc/articles/rcbc.html) for more details).

```{r}
# load package
library(rcbc)

# define optimization problem and solve it
## max 1 * x + 2 * y
## s.t.
##   x + y <= 1
##   x, y binary
result <- cbc_solve(
 obj = c(1, 2),
 mat = matrix(c(1, 1), ncol = 2, nrow = 1),
 is_integer = c(TRUE, TRUE),
 row_lb = -Inf, row_ub = 1, max = TRUE,
 col_lb = c(0, 0), col_ub = c(1, 1),
 cbc_args = list("SEC" = "1"))

# extract solution status
solution_status(result)

# extract solution values
column_solution(result)

# extract objective value for solution
objective_value(result)
```

## ROI plugin

There is now a work in progress [ROI plugin](https://github.com/dirkschumacher/ROI.plugin.cbc).

## Contribution

Feel free to open issues and send pull requests.

## Citation

Please cite the _rcbc_ R package and the [CBC solver](https://github.com/coin-or/Cbc) in publications.

```{r, echo  = FALSE, result = "asis", comment = ""}
citation("rcbc")
```
